Robust Trajectory Optimization on a Quantum System

### 2021-02-05 paper/0abstract.tex
jens edits 2 addressed or todo'd

%% ABSTRACT
\begin{abstract}
  The ability to engineer high-fidelity gates on quantum processors in the presence of
  systematic errors remains the primary challenge requisite to achieving quantum advantage.
  Quantum optimal control methods have proven effective in experimentally
  realizing high-fidelity gates, but they require exquisite calibration to be performant.
  We apply robust trajectory optimization techniques to suppress gate errors arising from system
  parameter uncertainty.
  We propose a method that takes advantage of uncertain parameter
  derivative information while maintaining
  computational efficiency by transforming high-order differential equations to coupled,
  first-order ODEs.
  Additionally, the effect of depolarization on a gate is most accurately modeled by
  integrating the Lindblad master equation,
  which is computationally expensive.
  We propose a computationally efficient model
  and utilize time-optimal control to achieve high fidelity gates in the presence of depolarization.
  We apply these techniques to a fluxonium qubit and suppress
  simulated gate errors due to parameter uncertainty below $10^{-7}$ for
  static parameter deviations on the order of $1\%$.
\end{abstract}

>>>>

%% ABSTRACT
\begin{abstract}
  The ability to engineer high-fidelity gates on quantum processors in the presence of
  systematic errors remains the primary challenge requisite to achieving quantum advantage.
  Quantum optimal control methods have proven effective in experimentally
  realizing high-fidelity gates, but they require exquisite calibration to be performant.
  We apply robust trajectory optimization techniques to suppress gate errors arising from system
  parameter uncertainty.
  We propose a derivative-based approach that maintains
  computational efficiency by using forward-mode differentiation.
  Additionally, the effect of depolarization on a gate is typically modeled by
  integrating the Lindblad master equation,
  which is computationally expensive.
  We employ a computationally efficient model
  and utilize time-optimal control to achieve high fidelity gates in the presence of depolarization.
  We apply these techniques to a fluxonium qubit and suppress
  simulated gate errors due to parameter uncertainty below $10^{-7}$ for
  static parameter deviations on the order of $1\%$.
\end{abstract}

### 2021-02-05 paper/1introduction.tex
jens edits 2 addressed or todo'd

In this work, we study three robust control techniques that
make the system's quantum state trajectory less sensitive
to static and time-dependent parameter uncertainty:
\begin{enumerate}
\item A sampling method, similar to the work of \cite{allen2019robust,
  ball2020software, carvalho2020error, khaneja2005optimal,
  reinhold2019controlling, rembold2020introduction}.
  \item An unscented sampling method adapted from the unscented transform used in
    state estimation \cite{howell2020direct, julier2004unscented,
      lee2013sigma, thangavel2020robust}.
  \item A derivative method, which penalizes the sensitivity of the quantum state trajectory
    to uncertain parameters.
\end{enumerate}
We apply these techniques to the fluxonium qubit presented in \cite{zhang2020universal}.
We also show that QOC can solve important problems associated with
fluxonium-based qubits: mitigating
depolarization by taking advantage of the $T_{1}$-dependence of the controls
and synchronizing qubits with distinct frequencies
by performing phase gates in arbitrary times.
To mitigate depolarization,
we perform time-optimal control and
employ an efficient depolarization model
for which the computational cost is independent of the
Hilbert space dimension.
Leveraging recent advances in trajectory optimization within the field of robotics, we
solve these optimization problems using ALTRO (Augmented Lagrangian TRajectory Optimizer)
\cite{howell2019altro}, which can enforce constraints on the quantum state trajectory
and control parameters.

>>>>

In this work, we study three robust control techniques that
make the system's quantum state trajectory less sensitive
to static and time-dependent parameter uncertainty:
\begin{enumerate}
\item A sampling method, similar to the work of \cite{allen2019robust,
  ball2020software, carvalho2020error, khaneja2005optimal,
  reinhold2019controlling, rembold2020introduction}.
  \item An unscented sampling method adapted from the unscented transform used in
    state estimation \cite{howell2020direct, julier2004unscented,
      lee2013sigma, thangavel2020robust}.
  \item A derivative method, which penalizes the sensitivity of the quantum state trajectory
    to uncertain parameters.
\end{enumerate}
We apply these techniques to the fluxonium qubit presented in \cite{zhang2020universal}.
We also show that QOC can solve important problems associated with
fluxonium-based qubits: exploiting the $T_{1}$-dependence of the controls
to mitigate depolarization
and synchronizing the phase of qubits with distinct frequencies.
To mitigate depolarization,
we perform time-optimal control and
employ an efficient depolarization model
for which the computational cost is independent of the
Hilbert space dimension.
Leveraging recent advances in trajectory optimization within the field of robotics, we
solve these optimization problems using ALTRO (Augmented Lagrangian TRajectory Optimizer)
\cite{howell2019altro}, which can enforce constraints on the quantum state trajectory
and control parameters.

### 2021-02-05 paper/2background.tex
jens edits 2 addressed or todo'd

\section{Background \label{sec:background}}
In this section, we
review the QOC problem statement
and describe the ALTRO solver \cite{howell2019altro}.
QOC concerns a vector of time-dependent control parameters $u(t)$
that steer the evolution of a (quantum) state $\ket{\psi(t)}$.
The evolution of the state is governed by the time-dependent
Schr{\"o}dinger equation (TDSE),
\begin{equation}
  i \hbar \frac{d}{dt} \ket{\psi(t)} = H(u(t), t) \ket{\psi(t)}.
  \label{eq:tdse}
\end{equation}
The Hamiltonian $H(u(t), t)$ is determined by the quantum system.
The QOC problem is to find the
controls that minimize a functional $J(u(t))$.
To make the problem numerically tractable,
the controls and state are discretized into $N$ knot points (time steps).
In the case of a single state transfer problem, the functional is
the infidelity of the desired final state and the initial state evolved
to the final knot point,
$J(u) = 1 - {\lvert \braket{\psi_{f}}{\psi_{N}(u)} \rvert}^{2}$.
In general, $J(u)$ is a linear combination of cost functions on the state
as well as cost functions on the controls.
Standard QOC solvers compute derivatives of the functional $\nabla_{u} J(u)$,
which can easily be used to implement first-order optimization methods
\cite{machnes2015tunable, khaneja2005optimal, leung2017speedup, goerz2019krotov}.

@@

ALM converges superlinearly, but poor numerical conditioning may lead
to small decreases in the constraint violations near the locally optimal solution
\cite{bertsekas1996constrained}.
To address this shortcoming, ALTRO
projects the solution from the ALM stage onto the constraint manifold using
a projected Newton method, achieving ultra-low
constraint violations $\sim 10^{-8}$.
For more information on the details of the ALTRO
solver, see \cite{howell2019altro, Jackson2020altroc}.

>>>>

\section{Background \label{sec:background}}
In this section, we
review the QOC problem statement
and describe the ALTRO solver \cite{howell2019altro}.
QOC concerns a vector of time-dependent control parameters $u(t)$
that steer the evolution of a (quantum) state $\ket{\psi(t)}$.
The evolution of the state is governed by the time-dependent
Schr{\"o}dinger equation (TDSE),
\begin{equation}
  i \hbar \frac{d}{dt} \ket{\psi(t)} = H(u(t), t) \ket{\psi(t)}.
  \label{eq:tdse}
\end{equation}
The Hamiltonian $H(u(t), t)$ is determined by the quantum system.
The QOC problem is to find the
controls that minimize a functional $J(u(t))$.
To make the problem numerically tractable,
the controls and state are discretized into $N$ knot points (time steps).
In the case of a single state-transfer problem, the functional is
the infidelity of the desired final state and the initial state evolved
to the final knot point,
$J(u) = 1 - {\lvert \braket{\psi_{f}}{\psi_{N}(u)} \rvert}^{2}$.
In general, $J(u)$ is a linear combination of cost functions on the state
as well as cost functions on the controls.
Standard QOC solvers compute derivatives of the functional $\nabla_{u} J(u)$,
which can easily be used to implement first-order optimization methods
\cite{machnes2015tunable, khaneja2005optimal, leung2017speedup, goerz2019krotov}.

@@

ALM converges superlinearly, but poor numerical conditioning may lead
to small decreases in the constraint violations near the locally optimal solution
\cite{bertsekas1996constrained}.
To address this shortcoming, ALTRO
projects the solution from the ALM stage onto the constraint manifold using
a (direct) projected Newton method, achieving ultra-low
constraint violations $\sim 10^{-8}$.
For more information on the details of the ALTRO
solver, see \cite{howell2019altro, Jackson2020altroc}.

### 2021-02-05 paper/3fluxonium.tex
jens edits 2 addressed or todo'd

First, we outline the constraints for the fluxonium gate problem.
We formulate this problem as a multi-state-transfer problem.
The initial conditions on
the states are $\ket{\psi^{0}_{1}} = \ket{0}$, $\ket{\psi^{1}_{1}} = \ket{1}$
\eqref{eq:istate_con}
where the superscript is an index $i \in \{0, 1\}$,
and the subscript indicates the first knot point $k = 1$.
The states at the final knot point are constrained to be
the image of the initial states under the desired gate $U$,
$\ket{\psi^{i}_{N}} = U \ket{\psi^{i}_{1}} \ \forall \ i$
\eqref{eq:tstate_con}.
Furthermore, we impose the normalization constraint
${\lvert \braket{\psi^{i}_{k}}{\psi^{i}_{k}} \rvert}^{2} = 1 \ \forall \ i,k$
\eqref{eq:statenorm_con}
to ensure the solver does not take advantage of discretization errors in numerical integration.
To refer to the discrete moments the flux, we introduce the notation
$\int^{t_{k}}_{t_{1}} a(t) \ \mathrm{d}t \equiv \int_{t} a_{k}$,
$a(t = t_{k}) \equiv a_{k}$,
$\pdv*[n]{a(t)}{t} \lvert_{t = t_{k}} \equiv \partial^{n}_{t} a_{k}$.
We impose the zero net flux constraint $\int_{t} a_{N} = 0$
\eqref{eq:znf_con}
which mitigates the inductive drift ubiquitous in flux-bias lines
\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
The flux is constrained by $\lvert a_{k} \rvert \leq 0.5 \ \textrm{GHz} \ \forall \ k$
\eqref{eq:amp_con}
to ensure the two-level approximation \eqref{eq:hamiltonian} remains valid.
We also enforce the boundary condition $a_{1} = a_{N} = 0$ \eqref{eq:bound_con}
so the gates may be concatenated arbitrarily. Additionally,
we have the initial condition $\int_{t} a_{1} = \partial_{t} a_{1} = 0$
\eqref{eq:ic_con}.

@@

  \begin{subfigure}{.315\textwidth}
    \includegraphics[width=\linewidth]{assets/f1a.png}
    \caption{\label{fig:longitudea}}
  \end{subfigure}\hfill
  \begin{subfigure}{.23\textwidth}
    \includegraphics[width=\linewidth]{assets/f1b.png}
    \caption{\label{fig:longitudeb}}
  \end{subfigure}\hfill
  \begin{subfigure}{.4\textwidth}
    \includegraphics[width=\linewidth]{assets/f1c.png}
    \caption{\label{fig:longitudec}}
  \end{subfigure}
  \caption{
    (a) Flux pulses for the numerical gates (dark blue)
    and the analytic gates (light pink).
    (b) $T_{1}$ interpolation function used in optimization. Circle markers
    indicate measured $T_{1}$ times. Non-circle markers
    indicate the time-averaged, absolute amplitude of each flux pulse.
    (c) Cumulative gate errors due to depolarization as a function of the
    number of gates applied.
    Cumulative gate errors for the numerical $Z/2$ and $Y/2$ gates
    are indistinguishable. Inset shows log-scaled cumulative gate errors
    for small gate counts.
  }
  \label{fig:longitude}
\end{figure*}

>>>>

First, we outline the constraints for the fluxonium gate problem.
All gates in this work satisfy these constraints to a tolerance
of $\sim 10^{-8}$.
The initial conditions on
the states are $\ket{\psi^{0}_{1}} = \ket{0}$, $\ket{\psi^{1}_{1}} = \ket{1}$
\eqref{eq:istate_con}
where the superscript is an index $i \in \{0, 1\}$,
and the subscript indicates the knot point $k = 1$.
The states at the final knot point are constrained to be
the image of the initial states under the desired gate $U$,
$\ket{\psi^{i}_{N}} = U \ket{\psi^{i}_{1}} \ \forall \ i$
\eqref{eq:tstate_con}.
Furthermore, we impose the normalization constraint
${\lvert \braket{\psi^{i}_{k}}{\psi^{i}_{k}} \rvert}^{2} = 1 \ \forall \ i,k$
\eqref{eq:statenorm_con}
to ensure the solver does not take advantage of discretization errors in numerical integration.
To refer to the discrete moments the flux, we introduce the notation
$\int^{t_{k}}_{t_{1}} a(t) \ \mathrm{d}t \equiv \int_{t} a_{k}$,
$a(t_{k}) \equiv a_{k}$,
$\pdv*[n]{a(t)}{t} \lvert_{t = t_{k}} \equiv \partial^{n}_{t} a_{k}$.
We impose the zero net flux constraint $\int_{t} a_{N} = 0$
\eqref{eq:znf_con}
which mitigates the inductive drift ubiquitous in flux-bias lines
\cite{battistel2019fast, krantz2019quantum, zhang2020universal}.
The flux is constrained by $\lvert a_{k} \rvert \leq 0.5 \ \textrm{GHz} \ \forall \ k$
\eqref{eq:amp_con}
to ensure the two-level approximation \eqref{eq:hamiltonian} remains valid.
We also enforce the boundary condition $a_{1} = a_{N} = 0$ \eqref{eq:bound_con}
so the gates may be concatenated arbitrarily. Additionally,
we have the initial condition $\int_{t} a_{1} = \partial_{t} a_{1} = 0$
\eqref{eq:ic_con}.

@@

  \begin{subfigure}{.315\textwidth}
    \includegraphics[width=\linewidth]{assets/f1a.png}
    \caption{\label{fig:longitudea}}
  \end{subfigure}\hfill
  \begin{subfigure}{.23\textwidth}
    \includegraphics[width=\linewidth]{assets/f1b.png}
    \caption{\label{fig:longitudeb}}
  \end{subfigure}\hfill
  \begin{subfigure}{.4\textwidth}
    \includegraphics[width=\linewidth]{assets/f1c.png}
    \caption{\label{fig:longitudec}}
  \end{subfigure}
  \caption{
    (a) Flux pulses for the numerical gates (dark blue)
    and the analytic gates (light pink).
    (b) $T_{1}$ interpolation function used in optimization. Circle markers
    indicate measured $T_{1}$ times. Non-circle markers
    are plotted at the time-averaged, absolute amplitude of each flux pulse.
    (c) Cumulative gate errors due to depolarization as a function of the
    number of gates applied.
    Cumulative gate errors for the numerical $Z/2$ and $Y/2$ gates
    are indistinguishable. Inset shows log-scaled cumulative gate errors
    for small gate counts.
  }
  \label{fig:longitude}
\end{figure*}

### 2021-02-05 paper/4longitude.tex
jens edits 2 addressed or todo'd

\section{Depolarization Mitigation\label{sec:longitude}}
In this section, we outline a method
for optimizing the flux to mitigate depolarization.
For many superconducting circuits, the 1/$e$ depolarization time ($T_{1}$) is independent of the control parameters,
so the fastest possible gate incurs the least depolarization error
\cite{schulteherbruggen2011optimal}.
For the fluxonium, however, $T_{1}$ is strongly dependent on the flux.
We enable the optimizer to trade longer gate times
for lower depolarizaiton times or vice-versa by making the gate time a decision variable.
Additionally, previous work has modeled the gate error due to depolarization
by evolving density matrices under a master
equation \cite{rembold2020introduction, schulteherbruggen2011optimal}
or evolving a large number of states in a quantum trajectory approach
\cite{abdelhafez2019gradient}.
We avoid the increase in computational complexity required for these
techniques by penalizing the integrated rate (probability) of depolarization in optimization.
Using this probability as a proxy for the gate error incurred
is reasonable because depolarization errors are incoherent and therefore
increase monotonically in time without interference.

The depolarization probability is given by,
\begin{equation}
  P_{1}(t) = \int_{0}^{t} T_{1}^{-1}(a(t^{\prime})) dt^{\prime}.
\end{equation}
This value is appended to the augmented state \eqref{eq:astatecontrols}
and its norm is penalized in the objective \eqref{eq:costfun} by setting
the corresponding element of the final augmented state to zero.
$T_{1}(a_{k})$ is obtained at each knot point by evaluating
a spline fit to experimental data of the form $\{(a, T_{1})\}$,
see Figure \ref{fig:longitudeb}.
It is also possible to fit a spline to theoretically obtained data.
However, experimental data often deviates from theoretical predictions
because it is difficult to account for all loss channels \cite{zhang2020universal},
and $T_{1}$ values often fluctuate on hour
time scales \cite{klimov2018fluctuations}.

We allow the optimizer to tune the gate time by
making the time step between each knot point $\Delta t_{k}$
a decision variable. Promoting $\Delta t_{k}$ to a decision variable, rather
than the number of knot points $N$, preserves the
Markovianity of the trajectory
optimization problem. The square root of the time step $\sqrt{\Delta t_{k}}$
is appended to the augmented control \eqref{eq:astatecontrols}
and the squared root
of the time step $\lvert \Delta t_{k} \rvert$ is used
for numerical integration in the discrete dynamics function \eqref{eq:dyn_con}.
To ensure numerical
integration accuracy is maintained, we constrain
the bounds of the time step at each knot point.

We analyze the effect of depolarization on
the $X/2$, $Y/2$, and $Z/2$ gates obtained with
our numerical method and the analytic gates. We
use the Lindblad master equation to simulate $T_{1}$ dissipation for successive
gate applications, and compute the cumulative gate error
after each application, see Appendix \ref{appendix:longitude}.
\todo{Jens, Danny, Ziwen, any idea what the correct way to simulate depolarization is
  when $T_{1}$ is time dependent?}.
The gate error reported in this text is the infidelity
of the evolved state and the target state averaged over 1000 pseudo-randomly
generated initial states.
The flux pulses for the numerical gates
are approximately periodic
with amplitudes $\sim 0.2 \textrm{GHz}$, see Figure \ref{fig:longitudea}.
They are reminiscent of the analytically determined Floquet operations
for a fluxonium described in \cite{huang2020engineering}
and realized in \cite{mundada2020floquet}.
The numerical gate times are greater
than the analytic gate times, but the
numerical flux pulses
spend more time at higher amplitudes, achieving higher $T_{1}$ times.
The single gate errors for both the analytic and numerical gates are
less than $10^{-4}$, which makes them sufficient for quantum error correction--a
prerequisite for fault-tolerant quantum
computing \cite{aharonov2008fault, knill2005quantum, gottesman1997stabilizer}.
However, the numerical gates achieve single gate errors
$\sim 5$ times less than those for the analytic gates,
which tracks closely with their relative improvement
on the depolarization probability metric, see Appendix \ref{appendix:longitude}.
This single gate error advantage corresponds to a significant reduction in error correction resources.
Furthermore, for successive gate applications, the gate error due to depolarization is approximately linear
in the gate count, which we expect for $t \ll T_{1}$, see Figure \ref{fig:longitudec}.
The gate error reduction for large gate counts is important for noisy, intermediate-scale quantum (NISQ)
applications. These improvements are significant for the constraints we have imposed on the gates,
and do not represent a fundamental limit to the optimization methods we have employed.

>>>>

\section{Depolarization Mitigation\label{sec:longitude}}
In this section, we outline a method
for optimizing the flux to mitigate depolarization.
For many superconducting circuits, the 1/$e$ depolarization time ($T_{1}$) is independent of the control parameters,
so the fastest possible gate incurs the least depolarization error
\cite{schulteherbruggen2011optimal}.
For the fluxonium, however, $T_{1}$ is strongly dependent on the flux.
We enable the optimizer to trade longer gate times
for lower depolarizaiton times or vice-versa by making the gate time a decision variable.
Additionally, previous work has modeled the gate error due to depolarization
by evolving density matrices under a master
equation \cite{rembold2020introduction, schulteherbruggen2011optimal}
or evolving a large number of states in a quantum trajectory approach
\cite{abdelhafez2019gradient}.
We avoid the increase in computational complexity required for these
techniques by penalizing the integrated depolarization rate in optimization.
Using this metric as a proxy for the gate error incurred
is reasonable because depolarization errors are incoherent and therefore
increase monotonically in time without interference.

The integrated depolarization rate is given by,
\begin{equation}
  D_{1}(t) = \int_{0}^{t} T_{1}^{-1}(a(t^{\prime})) dt^{\prime}.
\end{equation}
This value is appended to the augmented state \eqref{eq:astatecontrols}
and its norm is penalized in the objective \eqref{eq:costfun} by setting
the corresponding element of the final augmented state to zero.
$T_{1}(a_{k})$ is obtained at each knot point by evaluating
a spline fit to experimental data of the form $\{(a, T_{1})\}$,
see Figure \ref{fig:longitudeb}.
It is also possible to fit a spline to theoretically obtained data.
However, experimental data often deviates from theoretical predictions
because it is difficult to account for all loss channels \cite{zhang2020universal},
and $T_{1}$ values often fluctuate on hour
time scales \cite{klimov2018fluctuations}.

We allow the optimizer to tune the gate time by
making the time step between each knot point
a decision variable. 
The square root of the time step $\sqrt{\Delta t_{k}}$
is appended to the augmented control \eqref{eq:astatecontrols}
and the squared root
of the time step $\lvert \Delta t_{k} \rvert$ is used
for numerical integration in the discrete dynamics function \eqref{eq:dyn_con}.
To ensure numerical
integration accuracy is maintained, we constrain
the bounds of the time step at each knot point.
We promote the time steps
to decision variables, rather
than the number of knot points $N$, to ensure each cost function depends
only on the augmented state and augmented
controls at its own knot point $\ell_{k}(x_{k}, u_{k})$, which is necessary for iLQR.

We analyze the effect of depolarization on
the $X/2$, $Y/2$, and $Z/2$ gates obtained with
our numerical method and the analytic gates. We
use the Lindblad master equation to simulate $T_{1}$ dissipation for successive
gate applications, and compute the cumulative gate error
after each application, see Appendix \ref{appendix:longitude}.
The gate error reported in this text is the infidelity
of the evolved state and the target state averaged over 1000 pseudo-randomly
generated initial states.
The flux pulses for the numerical gates
are approximately periodic
with amplitudes $\sim 0.2 \textrm{GHz}$, see Figure \ref{fig:longitudea}.
They are reminiscent of the analytically determined Floquet operations
for a fluxonium described in \cite{huang2020engineering}
and realized in \cite{mundada2020floquet}.
The numerical gate times are greater
than the analytic gate times, but the
numerical flux pulses
spend more time at higher amplitudes, achieving higher $T_{1}$ times.
The single gate errors for both the analytic and numerical gates are
less than $10^{-4}$, which makes them sufficient for quantum error correction--a
prerequisite for fault-tolerant quantum
computing \cite{aharonov2008fault, knill2005quantum, gottesman1997stabilizer}.
However, the numerical gates achieve single gate errors
$\sim 5$ times less than those for the analytic gates,
which tracks closely with their relative improvement
on the integrated depolarization rate metric, see Appendix \ref{appendix:longitude}.
This single gate error advantage corresponds to a significant reduction in error correction resources.
Furthermore, for successive gate applications, the gate error due to depolarization is approximately linear
in the gate count, which we expect for $t \ll T_{1}$, see Figure \ref{fig:longitudec}.
The gate error reduction for large gate counts is important for noisy, intermediate-scale quantum (NISQ)
applications. These improvements are significant for the constraints we have imposed on the gates,
and do not represent a fundamental limit to the optimization methods we have employed.

### 2021-02-05 paper/5static.tex
jens edits 2 addressed or todo'd

\section{Robustness to Static Parameter Uncertainty \label{sec:static}}
We have formulated the QOC
problem as an open-loop optimization problem, equivalently,
we do not incorporate feedback from the experiment in optimization.
However, the device typically deviates from the Hamiltonian we use in optimization,
leading to poor experimental performance. We combat errors
of this form using robust control techniques,
making the state evolution insensitive
to Hamiltonian parameter uncertainty. As an example,
we mitigate errors arising from the drift and finite measurement
precision of the qubit frequency which modifies the fluxonium Hamiltonian
\eqref{eq:hamiltonian} by $f_{q} \rightarrow f_{q} + \delta f_{q}$.
We consider three robust control techniques to accomplish this task:
a sampling method, an unscented sampling method,
and a derivative method.

The sampling method incentivizes the optimizer
to ensure multiple instances of a state, each of which evolves
with a distinct value of the uncertain parameter, achieve
the same target state. Variants of this technique have been proposed
in the context of QOC
\cite{allen2019robust, ball2020software, khaneja2005optimal,
  reinhold2019controlling, rembold2020introduction} and applied
experimentally \cite{carvalho2020error}.
For each initial state,
we add two sample states $\ket{\psi^{\pm}}$
to the augmented state \eqref{eq:astatecontrols}. The discrete dynamics
function \eqref{eq:dyn_con} is modified
so the sample states evolve under the fluxonium Hamiltonian \eqref{eq:hamiltonian}
with $f_{q} \rightarrow f_{q} \pm \sigma_{f_{q}}$ for a fixed
hyperparameter $\sigma_{f_{q}}$ which is the standard deviation of the qubit frequency.
We penalize the infidelities of the sample states and their target state
by adding a cost function to the objective \eqref{eq:costfun} of the form
$\sum_{k, \pm} q_{k} (1 - {\lvert \braket{\psi^{\pm}_{k}}{U \psi_{1}} \rvert}^{2})$
where $q_{k}$ is a constant we supply.
For this method, the standard orthonormal basis states are an insufficient choice
for the initial states. As an example, a $Z/2$ gate achieved by idling
at the flux frustration point $a_{k} = 0 \ \forall \ k$
will be robust to qubit frequency detunings for the initial states $\ket{0}$
or $\ket{1}$ because the infidelity metric is insensitive to global phases,
but this gate will not be robust for any other initial states.
Therefore, we choose four initial states
so that their outer products
span the operators on the Hilbert space
$\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized},
which we refer to as the operator basis.

Whereas the sampling method penalizes the deviations of the sample states
from the target state, the unscented sampling method
penalizes the deviations of the sample states from the nominal state
\cite{howell2020direct, lee2013sigma,
  thangavel2020robust}. Accordingly, the cost function we add
to the objective \eqref{eq:costfun} takes the form
$\sum_{k, j} q_{k} (\psi^{j}_{k} - \psi_{k})^{T}
(\psi^{j}_{k} - \psi_{k})$, where $q_{k}$ is a
constant we supply, $\psi_{k}$ is
the evolved initial state (nominal state), and $\psi^{j}_{k}$ is a sample state
that evolves under a modified Hamiltonian similar to that in the sampling method.
We omit bra-ket notation here to emphasize that
the states are real vectors, and are given by the right-hand-side of the
isomorphism \eqref{eq:isomorphism}. Additionally,
the unscented sampling method employs the unscented transform
\cite{julier2004unscented, uhlmann1995dynamic}
to prevent the sample states
from drifting too far or close to the nominal state during evolution,
which would result in exceedingly large or small deviations in the cost function.
In particular, the sample states are chosen to encode a unimodal distribution over
the $2n$ elements of the nominal state, modeling the uncertainty in the state
as a result of the uncertainty in the parameter. The unscented transform
accurately propagates the mean and covariance of this distribution between
knot points, or equivalently, through the transformation of the TDSE.
For this method, we sample on one initial state $(\ket{0} + i\ket{1}) / \sqrt{2}$,
and do not observe a performance increase
using more initial states, for example those from the operator basis.
A detailed procedure for the unscented transformation is given
in Appendix \ref{appendix:unscented}.

The derivative method penalizes the sensitivity of the state
to the uncertain parameter, which is encoded in the $l$\textsuperscript{th}-order
state derivative $\partial_{f_{q}}^{l} \ket{\psi}$. In the $m$\textsuperscript{th}-order
derivative method, we append all state derivatives of order $1, \dots, m$
to the augmented state vector \eqref{eq:astatecontrols}
for each initial state. We use the initial state $\ket{0}$ for this method,
and observe no advantage to using more initial states.
We penalize the norms of the state derivatives
in the objective \eqref{eq:costfun} by setting the corresponding elements
of the final augmented state to zero.
We could obtain the state derivatives at each knot point
with backward-mode differentiation.
In a naive automatic differentiation scheme,
the discrete dynamics function at knot points
$1, \dots, k - 1$ would be differentiated to obtain the state
derivative at knot point $k$, requiring
$O(N^{2})$ matrix multiplications. Instead, we 
employ forward-mode differentiation on the TDSE \eqref{eq:tdse}
to obtain coupled, first-order ODEs
which require $O(N)$ matrix multiplications to integrate.
For example, the dynamics for the $1$\textsuperscript{st}-order derivative method are:
\begin{align}
  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi},\\
  i \hbar \frac{d}{dt} \ket{\partial_{f_{q}}\psi} &=
  H \ket{\partial_{f_{q}} \psi} +
  (\partial_{f_{q}} H) \ket{\psi}.
  \label{eq:d1dyn}
\end{align}
We integrate the coupled ODEs with exponential
integrators, see Appendix \ref{appendix:derivative},
in the discrete dynamics function \eqref{eq:dyn_con}. For runtimes
and complexity analyses for the three robust control techniques,
consult Appendix \ref{appendix:time}.

>>>>

\section{Robustness to Static Parameter Uncertainty \label{sec:static}}
We have formulated the QOC
problem as an open-loop optimization problem, equivalently,
we do not incorporate feedback from the experiment in optimization.
However, the device's parameters deviate from the parameters we use in optimization,
leading to poor experimental performance. We combat errors
of this form using robust control techniques,
making the state evolution insensitive
to parameter uncertainty. As an example,
we mitigate errors arising from the drift and finite measurement
precision of the qubit frequency which modifies the fluxonium Hamiltonian
\eqref{eq:hamiltonian} by $f_{q} \rightarrow f_{q} + \delta f_{q}$.
We consider three robust control techniques to accomplish this task:
a sampling method, an unscented sampling method,
and a derivative method.

The sampling method incentivizes the optimizer
to ensure multiple copies of a state, each of which evolves
with a distinct value of the uncertain parameter, achieve
the same target state. Variants of this technique have been proposed
in the context of QOC
\cite{allen2019robust, ball2020software, khaneja2005optimal,
  reinhold2019controlling, rembold2020introduction} and applied
experimentally \cite{carvalho2020error}.
For each initial state,
we add two sample states $\ket{\psi^{\pm}}$
to the augmented state \eqref{eq:astatecontrols}. The discrete dynamics
function \eqref{eq:dyn_con} is modified
so the sample states evolve under the fluxonium Hamiltonian \eqref{eq:hamiltonian}
with $f_{q} \rightarrow f_{q} \pm \sigma_{f_{q}}$ for a fixed
hyperparameter $\sigma_{f_{q}}$ which is the standard deviation of the qubit frequency.
We penalize the infidelities of the sample states and their target state
by adding a cost function to the objective \eqref{eq:costfun} of the form
$\sum_{k, \pm} b_{k} (1 - {\lvert \bra*{\psi^{\pm}_{k}}U\ket*{\psi_{1}} \rvert}^{2})$
where $b_{k}$ is a constant we supply.
For this method, the standard orthonormal basis states are an insufficient choice
for the initial states. As an example, a $Z/2$ gate achieved by idling
at the flux frustration point $a_{k} = 0 \ \forall \ k$
will be robust to qubit frequency detunings for the initial states $\ket{0}$
or $\ket{1}$ because the infidelity metric is insensitive to global phases,
but this gate will not be robust for any other initial states.
Therefore, we choose four initial states
so that their outer products
span the operators on the Hilbert space
$\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
(\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized},
which we refer to as the operator basis.

Whereas the sampling method penalizes the deviations of the sample states
from the target state, the unscented sampling method
penalizes the deviations of the sample states from the nominal state
\cite{howell2020direct, lee2013sigma,
  thangavel2020robust}. Accordingly, the cost function we add
to the objective \eqref{eq:costfun} takes the form
$\sum_{k, j} c_{k} (\psi^{j}_{k} - \psi_{k})^{T}
(\psi^{j}_{k} - \psi_{k})$, where $c_{k}$ is a
constant we supply, $\psi_{k}$ is
the evolved initial state (nominal state), and $\psi^{j}_{k}$ is a sample state
that evolves under a modified Hamiltonian similar to that in the sampling method.
We omit bra-ket notation here to emphasize that
the states are real vectors, and are given by the right-hand-side of the
isomorphism \eqref{eq:isomorphism}. Additionally,
the unscented sampling method employs the unscented transform
\cite{julier2004unscented, uhlmann1995dynamic}
to prevent the sample states
from drifting too far or close to the nominal state during evolution,
which would result in exceedingly large or small deviations in the cost function.
In particular, the sample states are chosen to encode a unimodal distribution over
the $2n$ elements of the nominal state, modeling the uncertainty in the state
as a result of the uncertainty in the parameter. The unscented transform
accurately propagates the mean and covariance of this distribution between
knot points, or equivalently, through the transformation of the TDSE.
For this method, we sample on one initial state $(\ket{0} + i\ket{1}) / \sqrt{2}$,
and do not observe a performance increase
using more initial states, for example those from the operator basis.
A detailed procedure for the unscented transformation is given
in Appendix \ref{appendix:unscented}.

The derivative method penalizes the sensitivity of the state
to the uncertain parameter, which is encoded in the $l$\textsuperscript{th}-order
state derivative $\partial_{f_{q}}^{l} \ket{\psi}$. In the $m$\textsuperscript{th}-order
derivative method, we append all state derivatives of order $1, \dots, m$
to the augmented state vector \eqref{eq:astatecontrols}
for each initial state. We use the initial state $\ket{0}$ for this method,
and observe no advantage to using more initial states.
We penalize the norms of the state derivatives
in the objective \eqref{eq:costfun} by setting the corresponding elements
of the final augmented state to zero.
We could obtain the state derivatives at each knot point
with backward-mode differentiation.
In a naive automatic differentiation scheme,
the discrete dynamics function at knot points
$1, \dots, k - 1$ would be differentiated to obtain the state
derivative at knot point $k$, requiring
$O(N^{2})$ matrix multiplications. Instead, we 
employ forward-mode differentiation on the TDSE \eqref{eq:tdse}
to obtain coupled, first-order ODEs
which require $O(N)$ matrix multiplications to integrate.
For example, the dynamics for the $1$\textsuperscript{st}-order derivative method are:
\begin{align}
  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi},\\
  i \hbar \frac{d}{dt} \ket{\partial_{f_{q}}\psi} &=
  H \ket{\partial_{f_{q}} \psi} +
  (\partial_{f_{q}} H) \ket{\psi}.
  \label{eq:d1dyn}
\end{align}
We integrate the coupled ODEs with exponential
integrators, see Appendix \ref{appendix:derivative},
in the discrete dynamics function \eqref{eq:dyn_con}. For runtimes
of the three robust control techniques,
consult Appendix \ref{appendix:time}.
